---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uncrtnty

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of {uncrtnty} is to provide functions to deal with uncertainty in NONMEM analyses.

  * Parse information about the "Block Form" of an OMEGA/SIGMA matrix from:
    + a *.lst* file with `parse_blockform_from_lst()`.
    + the matrix itself with `infer_blockform()`.
  * Turn a full OMEGA/SIGMA matrix into a list of OMEGA/SIGMA matrix blocks with `matrix_to_list()`.
  * Parse NONMEM estimate file (*.ext*) and covariance matrix file (*.cov*) into convenient output-list with `parse_ext()` and `parse_cov()`.
  * Compute degrees of freedom of the Inverse-Wishart distribution for OMEGA/SIGMA matrices with `compute_df()`.

These functions are executed under the hood in the following functions, which turns an [xpose](https://github.com/UUPharmacometrics/xpose) database object into:

  * a list of arguments to simulate with uncertainty, for instance with [simpar](https://github.com/metrumresearchgroup/simpar) with `xpose_to_simpar()`.
  * the code needed for the "$PRIOR NWPRI" routine in NONMEM (*yet to come... soon*).

## Installation

You can install the development version of uncrtnty from [github](https://github.com/FelicienLL/uncrtnty) with:

```{undefined eval=FALSE}
install.packages("remotes")
remotes::install_github("FelicienLL/uncrtnty")
```

## Example

```{r}
library(uncrtnty)
x <- readRDS(system.file("xposerun", "xpdb_ex_pk.rds", package = "uncrtnty"))
#Example NONMEM run from the xpose package
xpdb_to_simpar(x)
```

