---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uncrtnty

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

{uncrtnty} provides functions and an "uncrtnty"-list class in order to work with uncertainty in NONMEM analyses.

## Installation

You can install the development version of uncrtnty from [github](https://github.com/FelicienLL/uncrtnty) with:

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("FelicienLL/uncrtnty")
```

## The big idea

In NONMEM analyses, the uncertainty about parameter estimation can be described as followed. The estimation of THETA follows a multivariate normal distribution, with a mean vector (`th_est`) and a covariance matrix (`th_unc`) that carries the information about uncertainty. The estimates of OMEGA/SIGMA are matrices (`om_est/si_est`), and the uncertainty is informed by degrees of freedom (`om_unc/si_unc`). In {uncrtnty}, these information are storred into a single "uncrtnty"-list object.


```{r}
library(uncrtnty)
u_example
```

This "uncrtnty"-list object can be created:

  * from an [xpose](https://github.com/UUPharmacometrics/xpose) database object: `u_from_xpdb()`.
  * from NONMEM output files (*.lst*, *.ext*, *.cov*): *available soon*.
  * from a NONMEM *.xml* output file: *available soon*.
  * from a model reported in the literature: *available soon*

This "uncrtnty"-list object can be used to generate: 

  * the code implemented within the NONMEM $PRIOR NWPRI routine: `u_to_nwpri()`
  * the arguments to simulate with uncertainty thanks to the [simpar](https://github.com/metrumresearchgroup/simpar) package: `u_to_simpar()`

## Applications
#### Create the code for the $PRIOR routine in NONMEM
```{r}
u_to_nwpri(u_example)
```

#### Create the arguments to simulate with uncrtnty in simpar
```{r}
u_to_simpar(u_example, nsim = 100)
```


## Other useful features

```{r include=FALSE}
lstfile <- readLines(system.file("nm", "run003a.lst", package = "uncrtnty"))
m <- matrix(c(1, 0.12, 0.13, 0, 0, 0.12, 2, 0.23, 0, 0, 0.13, 0.23, 3, 0, 0, 0, 0, 0, 4, 0, 0, 0, 0 ,0 ,5), nrow = 5)

```




  * Parse information about the "Block Form" of an OMEGA/SIGMA matrix from:
    + a *.lst* file with `parse_blockform_from_lst()`.
```{r}
cat(lstfile[84:110], sep = "\n") # zoom inside a NONMEM .lst file
parse_lst(lst = lstfile)$om_blockform
parse_lst(lst = lstfile)$si_blockform
```
  *   
    + the matrix itself with `infer_blockform()`.
    
```{r}
m
infer_blockform(m)
```
  
  * Turn a full OMEGA/SIGMA matrix into a list of OMEGA/SIGMA matrix blocks with `matrix_to_list()`.
  
```{r}
matrix_to_list(m)
```
  
  * Parse NONMEM estimate file (*.ext*) and covariance matrix file (*.cov*) into convenient output-list with `parse_ext()` and `parse_cov()`.
  
  * Compute degrees of freedom of the Inverse-Wishart distribution for OMEGA/SIGMA matrices with `compute_df()`.
```{r}
est_om <- matrix(c(0.2, 0.01, 0.01, 0.1), ncol = 2)
se_om <- matrix(c(0.02, 0.005, 0.005, 0.03), ncol = 2)
compute_df(est = est_om, se = se_om)
```
  
  * Parse the covariance matrix of estimations of ETA in the context on individual parameter estimation `parse_phi()`.
